### 2024-03-04 content/chapter_3.tex
:construction: Add section titles

\chapter{Getting started}

\begin{section}{Introduction}
	% The quest for reliable, high-speed, and ubiquitous wireless connectivity has been long-standing since Marconi's illuminating radio in 1895.
	% Great successes have been made at transmitter and receiver sides over the past century, and the society is unprecedentedly close to the Shannon limit \cite{Shannon1948}.
	Today we are witnessing a paradigm shift from connectivity to intelligence, where the wireless environment is no longer a chaotic medium but a conscious agent that serves on demand.
	This is empowered by the recent advances in \gls{ris}, a real-time programmable metasurface of numerous non-resonant sub-wavelength scattering elements.
	It can manipulate the amplitude, phase, frequency, and polarization of the scattered waves \cite{Basar2019} with a higher energy efficiency, lower cost, lighter footprint, and greater scalability than relays.
	Using \gls{ris} for {passive beamforming} has attracted significant interest in wireless communication \cite{Wu2019,Wu2020c,Yang2020,Zheng2021}, backscatter \cite{Jia2020,Liang2022}, sensing \cite{Liu2022a,Hua2023}, and power transfer literature \cite{Wu2021d,Feng2022,Zhao2022}, reporting a second-order array gain and fourth-order power scaling law (with proper waveform).
	On the other hand, \gls{ris} also enables {backscatter modulation} by dynamically switching between different patterns, as already investigated \cite{Karasik2020,Basar2020,Zhao2022a} and prototyped \cite{Tang2019b,Dai2020a}.
	Despite fruitful outcomes, one critical unanswered question is the {channel shaping} capability: \emph{To what extent can a passive \gls{ris} reshape the wireless channel?}

	The answer indeed depends on the hardware architecture and scattering model.
	In conventional (a.k.a. diagonal) \gls{ris}, each scattering element is tuned by a dedicated impedance and acts as an \emph{individual} phase shifter \cite{Wu2020}.
	The concept is generalized to \gls{bd}-\gls{ris} \cite{Shen2020a,Li2023b} which groups adjacent elements using passive components.
	This allows \emph{cooperative} scattering --- wave impinging on one element can propagate within the circuit and depart partially from any element in the same group.
	\gls{bd}-\gls{ris} can thus control both amplitude and phase of the reflected wave, generalizing the scattering matrix from diagonal with unit-magnitude entries to block diagonal with  unitary blocks.
	Its benefit has been recently shown in receive power maximization \cite{Nerini2023,Santamaria2023,Fang2023,Nerini2023a}, transmit power minimization \cite{Zhou2023}, and rate maximization \cite{Zhou2023,Nerini2023a,Li2023d,Bartoli2023,Li2023c}.
	Practical issues such as channel estimation \cite{Li2023e} and mutual coupling \cite{Li2023f} have also been investigated.
	Therefore, \gls{bd}-\gls{ris} is envisioned as the next generation channel shaper with stronger signal processing flexibility \cite{Li2023g}.

	% Attempts to characterize the channel shaping capability can be classified into \emph{singular value centric} and \emph{power centric} methods.
	% Channel shaping is different from passive beamforming as it seeks to modify the inherent properties of the channel itself, allowing one to decouple \gls{ris} and transceiver design.
	Channel shaping is different from passive beamforming as it seeks to modify the inherent properties of the channel itself.
	This allows one to decouple the \gls{ris}-transceiver design and explore the fundamental limits of channel manipulation.
	For example, diagonal \gls{ris} has been proved useful for improving channel power \cite{Ning2020}, degree of freedom \cite{Ozdogan2020a,Li2023h}, condition number \cite{Zheng2022,Huang2023}, and effective rank \cite{ElMossallamy2021,Meng2023} in \gls{mimo}.
	In contrast, \gls{bd}-\gls{ris} can provide a higher channel power but existing results are limited to \gls{siso}\footnote{In terms of channel shaping, single-stream \gls{mimo} with given precoder and combiner \cite{Nerini2023} is equivalent to \gls{siso}.}. \cite{Nerini2023} and \gls{miso} \cite{Santamaria2023}.
	While these studies offer promising glimpses into the channel shaping potential, a comprehensive understanding of the capabilities and limitations is desired, and a universal design framework is missing.
	This paper aims to answer the channel shaping question through theoretical analysis and numerical optimization.
	The contributions are summarized below.

	First, we quantify the capability of a \gls{bd}-\gls{ris} to reshape the \gls{mimo} point-to-point channel in terms of singular values.
	The \emph{Pareto frontiers} are characterized by optimizing the {weighted sum of singular values}, where the weights can be positive, zero, or negative.
	% This generalizes most singular value metrics and provides a powerful design framework.
	The resulting singular value region generalizes most relevant metrics and provides an intuitive channel shaping benchmark.
	We then discuss some analytical singular value bounds in rank-deficient and fully-connected scenarios, which help to demystify the gain from off-diagonal entries.
	This is the first paper to answer the channel shaping question and highlight the \gls{bd}-\gls{ris} gain from a Pareto perspective.

	Second, we propose a \gls{rcg} algorithm adapted from \cite{Abrudan2008,Abrudan2009} for smooth optimization problems of asymmetric \gls{bd}-\gls{ris} with arbitrary group size.
	Specifically, block-wise update is performed along the geodesics\footnote{A geodesic refers to the shortest path between two points in a Riemannian manifold.} of the Stiefel manifold, which are expressed compactly by the exponential map \cite{Edelman1998}.
	It features lower complexity and faster convergence than general manifold optimization \cite{Absil2009,Pan2022d}, and solves the Pareto singular value problem.
	This is the first paper to tailor an efficient optimization framework for asymmetric \gls{bd}-\gls{ris}.

	Third, we tackle \gls{bd}-\gls{ris} \gls{mimo} rate maximization with two solutions: a local-optimal approach through \gls{ao} and a low-complexity approach over channel shaping.
	The former updates active and passive beamformers by eigenmode transmission and \gls{rcg} algorithm, respectively.
	The latter suboptimally decouples both blocks, recasts the shaping problem as channel power maximization, and solves it iteratively in closed form.
	Interestingly, the gap in between vanishes as \gls{bd}-\gls{ris} evolves from diagonal (single-connected) to unitary (fully-connected).
	It suggests channel shaping offers a promising low-complexity solution for joint \gls{ris}-transceiver designs.

	% Third, we propose a closed-form iterative algorithm for power centric channel shaping problems.
	% The idea is to successively approximate the quadratic objective by a sequence of affines and solve the local problems by \gls{svd}.
	% Case studies are conducted for channel power maximization in \gls{pc} and leakage interference minimization in \gls{ic}.

	Fourth, extensive simulations reveal that the performance gain from \gls{bd}-\gls{ris} increases with group size and \gls{mimo} dimensions.
	In terms of channel power, fully-connected \gls{bd}-\gls{ris} boosts up to 62\% and 270\% over diagonal \gls{ris} in $1 \times 1$ and $4 \times 4$ \gls{mimo} under independent Rayleigh fading, respectively.
	The superiority stems from stronger \emph{subchannel rearrangement} and \emph{subspace alignment} capabilities empowered by in-group cooperation.
	It emphasizes the importance of using \gls{bd}-\gls{ris} in large-scale \gls{mimo} systems.

\end{section}

\begin{section}{\glsfmtlong{bd}-\glsfmtlong{ris} Model}
	Consider a \gls{bd}-\gls{ris} aided point-to-point \gls{mimo} system with $N_\mathrm{T}$, $N_\mathrm{S}$, $N_\mathrm{R}$ transmit, scatter, and receive antennas, respectively.
	This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$ in the following context.
	The \gls{bd}-\gls{ris} can be modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that divides into $G$ individual groups, each containing $L \triangleq N_\mathrm{S} / G$ elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
	To simplify the analysis, we assume a lossless asymmetric network\footnote{While symmetric \gls{bd}-\gls{ris} (involving capacitors and inductors) is often considered in the literature \cite{Shen2020a,Nerini2023,Santamaria2023,Fang2023,Nerini2023a,Zhou2023,Li2023d,Bartoli2023}, asymmetric \gls{bd}-\gls{ris} can be built over reconfigurable asymmetric passive components (e.g., ring hybrids and branch-line hybrids) \cite{Ahn2006}.} without mutual coupling between scatter elements, as previously considered in \cite{Li2023b,Li2023c,Bartoli2023}.
	The overall scattering matrix of the \gls{bd}-\gls{ris} is
	\begin{equation}
		\mathbf{\Theta} = \mathrm{diag}(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),
	\end{equation}
	where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is a unitary matrix that describes the response of group $g \in \mathcal{G} \triangleq \{1, \ldots, G\}$.
	That is, $\mathbf{\Theta}_g = \mathbf{\Theta}_g^\mathsf{H}$ and $\mathbf{\Theta}_g^\mathsf{H} \mathbf{\Theta}_g = \mathbf{I}$.
	Note that diagonal (single-connected) and unitary (fully-connected) \gls{ris} can be regarded as its extreme cases with group size $L=1$ and $L=N_\mathrm{S}$, respectively.
	Some potential physical architectures of \gls{bd}-\gls{ris} are illustrated in \cite[Fig. 3]{Shen2020a}, \cite[Fig. 5]{Li2023c}, and \cite[Fig. 2]{Nerini2023a}, where the radiation pattern and circuit topology have impacts on the scattering matrix.

	Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct (transmitter-receiver), backward (\gls{ris}-receiver), and forward (transmitter-\gls{ris}) channels, respectively.
	The equivalent channel is a function of the scattering matrix
	\begin{equation}
		\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
		\label{eq:channel_equivalent}
	\end{equation}
	where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward subchannels for \gls{ris} group $g$, corresponding to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$, respectively.
	% That is, $\mathbf{H}_{\mathrm{B},g}$ (resp. $\mathbf{H}_{\mathrm{F},g}$) corresponds to the $(g-1)L$ to $gL$ columns (resp. rows) of $\mathbf{H}_\mathrm{B}$ (resp. $\mathbf{H}_\mathrm{F}$).
	% , corresponding to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$.
	% $\mathbf{H}_{\mathrm{B},g}$ (resp. $\mathbf{H}_{\mathrm{F},g}$) corresponds to the $(g-1)L$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$
	Since unitary matrices constitute an algebraic group with respect to multiplication, the scattering matrix of group $g$ can be decomposed as
	\begin{equation}
		\mathbf{\Theta}_g = \mathbf{L}_g \mathbf{R}_g^\mathsf{H},
	\end{equation}
	where $\mathbf{L}_g, \mathbf{R}_g \in \mathbb{U}^{L \times L}$ are two unitary factor matrices.
	Let $\mathbf{H}_{\mathrm{B},g} = \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \mathbf{V}_{\mathrm{B},g}^\mathsf{H}$ and $\mathbf{H}_{\mathrm{F},g} = \mathbf{U}_{\mathrm{F},g} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}$ be the compact \gls{svd} of the backward and forward channels, respectively.
	The equivalent channel can thus be rewritten as
	\begin{equation}
		\mathbf{H} = \overbrace{\mathbf{H}_\mathrm{D} + \sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g \mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_\text{backward-forward} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}}^\text{direct-indirect}.
		\label{eq:channel_equivalent_svd}
	\end{equation}

	\begin{remark}
		By analyzing \eqref{eq:channel_equivalent_svd}, we conclude that channel shaping through \gls{bd}-\gls{ris} creates two critical opportunities:
		\begin{itemize}
			\item \textbf{Subspace alignment:} each group can align the singular matrices of the backward-forward (intra-group, multiplicative) and the direct-indirect (inter-group, additive) channels. In \gls{siso}, subspace alignment boils down to phase matching and the optimal scattering matrix of group $g$ is
			\begin{equation}
				\mathbf{\Theta}_g^\star = \exp \bigl(\jmath \mathrm{arg}(h_\mathrm{D})\bigr) \mathbf{V}_{\mathrm{B},g} \mathbf{U}_{\mathrm{F},g}^\mathsf{H},

>>>>

\chapter{\glsfmtshort{ris}-Aided \glsfmtshort{swipt}: Joint Waveform and Beamforming Design}\label{ch:ris_aided_swipt}

\begin{section}{Introduction}
	\begin{subsection}{\glsfmtlong{swipt}}
		With the great advance in communication performance, a bottleneck of wireless networks has come to energy supply. \gls{swipt} is a promising solution to connect and power mobile devices via \gls{rf} waves. It provides low power at \si{\uW} level but broad coverage up to hundreds of meters in a sustainable and controllable manner, bringing more opportunities to the \gls{iot} and \gls{m2m} networks. The upsurge in wireless devices, together with the decrease of electronics power consumption, calls for a re-thinking of future wireless networks based on \gls{wpt} and \gls{swipt} \cite{Clerckx2019}.

		The concept of \gls{swipt} was first cast in \cite{Varshney2008}, where the authors investigated the \gls{r-e} tradeoff for a flat Gaussian channel and typical discrete channels. \cite{Zhou2013} proposed two practical co-located information and power receivers, i.e., \gls{ts} and \gls{ps}. Dedicated information and energy beamforming were then investigated in \cite{Zhang2013,Park2014} to characterize the \gls{r-e} region for multi-antenna broadcast and interference channels. On the other hand, \cite{Trotter2009} pointed out that the \gls{rf}-to-\gls{dc} conversion efficiency of rectifiers depends on the input power and waveform shape. It implies that the modeling of the energy harvester, particularly its nonlinearity, has a crucial impact on the waveform preference, resource allocation, and system design of any wireless-powered systems \cite{Trotter2009,Clerckx2018,Clerckx2019}. Motivated by this, \cite{Clerckx2016a} derived a tractable nonlinear harvester model based on the Taylor expansion of diode I-V characteristics, and performed joint waveform and beamforming design for \gls{wpt}. Simulation and experiments showed the benefit of modeling energy harvester nonlinearity in real system design \cite{Kim2019,Kim2020a} and demonstrated the joint waveform and beamforming strategy as a key technique to expand the operation range \cite{Kim2021}. A low-complexity adaptive waveform design by \gls{smf} was proposed in \cite{Clerckx2017} to exploit the rectifier nonlinearity, whose advantage was then demonstrated in a prototype with channel acquisition \cite{Kim2017}. Beyond \gls{wpt}, \cite{Clerckx2018b} uniquely showed that the rectifier nonlinearity brings radical changes to \gls{swipt} design, namely (i) modulated and unmodulated waveforms are not equally suitable for wireless power delivery; (ii) a multi-carrier unmodulated waveform superposed to a multi-carrier modulated waveform can enlarge the \gls{r-e} region; (iii) a combination of \gls{ps} and \gls{ts} is generally the best strategy; (iv) the optimal input distribution is not the conventional \gls{cscg}; (v) modeling rectifier nonlinearity is beneficial to system performance and essential to efficient \gls{swipt} design. Those observations, validated experimentally in \cite{Kim2019}, led to the question: \emph{What is the optimal input distribution for \gls{swipt} under nonlinearity?} This question was answered in \cite{Varasteh2020} for single-carrier \gls{swipt}, and some attempts were further made in \cite{Varasteh2019d} for multi-carrier \gls{swipt}. The answers shed new light to the fundamental limits of \gls{swipt} and practical signaling (e.g., modulation and waveform) strategies. It is now well understood from \cite{Clerckx2018b,Varasteh2020,Varasteh2019d} that, due to harvester nonlinearity, a combination of \gls{cscg} and on-off keying in single-carrier setting and non-zero mean asymmetric inputs in multi-carrier setting lead to significantly larger \gls{r-e} region compared to conventional \gls{cscg}. Recently, \cite{Varasteh2020a} used machine learning techniques to design \gls{swipt} signaling under nonlinearity to complement the information-theoretic results of \cite{Varasteh2020}, and new modulation schemes were subsequently invented.
	\end{subsection}

	\begin{subsection}{\glsfmtlong{ris}}
		\gls{ris} has recently emerged as a promising technique that adapts the propagation environment to enhance the spectrum and energy efficiency. In practice, a \gls{ris} consists of multiple individual sub-wavelength reflecting elements to adjust the amplitude and phase of the incoming signal (i.e., passive beamforming). Different from the relay, backscatter and frequency-selective surface \cite{Anwar2018}, the \gls{ris} assists the primary transmission using passive components with negligible thermal noise but is limited to frequency-dependent reflection.

		Inspired by the development of real-time reconfigurable metamaterials \cite{Cui2014}, the authors of \cite{Liaskos2018} introduced a programmable metasurface that steers or polarizes the electromagnetic wave at a specific frequency to mitigate signal attenuation. \cite{Wu2018} proposed a \gls{ris}-assisted \gls{miso} system and jointly optimized the precoder at the \gls{ap} and the phase shifts at the \gls{ris} to minimize the transmit power. The active and passive beamforming problem was then extended to the discrete phase shift case \cite{Wu2019a} and the multi-user case \cite{Wu2019}. In \cite{Abeywickrama2020}, the authors investigated the impact of non-zero resistance on the reflection pattern and emphasized the coupling between reflection amplitude and phase shift in practice. To estimate the cascaded \gls{ap}-\gls{ris}-\gls{ue} link without \gls{rf}-chains at the \gls{ris}, practical protocols were developed based on element-wise on/off switching \cite{Nadeem2019}, training sequence and reflection pattern design \cite{You2019,Kang2020}, and compressed sensing \cite{Wang2020}. The hardware architecture, design challenges, and application opportunities of practical \gls{ris} were covered in \cite{Wu2020}. In \cite{Dai2020}, a prototype \gls{ris} with \num{256} \num{2}-bit elements based on \gls{pin} diodes was developed to support real-time video transmission at \si{GHz} and mmWave frequency.
	\end{subsection}

	\begin{subsection}{\glsfmtshort{ris}-Aided \glsfmtshort{swipt}}
		By integrating \gls{ris} with \gls{swipt}, the constructive reflection can boost the end-to-end power efficiency and improve the \gls{r-e} tradeoff. In multi-user cases, dedicated energy beams were proved unnecessary for the weighted sum-power maximization \cite{Wu2020b} but essential when fairness issue is considered \cite{Tang2019}. It was also claimed that \gls{los} links could boost the power efficiency since rank-deficient channels require fewer energy beams \cite{Wu2020a}. However, \cite{Wu2020b,Tang2019,Wu2020a} were based on a linear energy harvester model that is known in both the \gls{rf} and the communication literature to be inefficient and inaccurate \cite{Clerckx2019,Trotter2009,Clerckx2018,Clerckx2016a,Kim2019,Kim2020a,Kim2021,Clerckx2017,Kim2017,Clerckx2018b,Varasteh2020,Varasteh2019d,Varasteh2020a}. Based on practical \gls{ris} and harvester models, \cite{Xu2021c} introduced a scalable resource allocation framework for a large-scale tile-based \gls{ris}-assisted \gls{swipt} system, where the optimization consists of a reflection design stage and a joint reflection selection and precoder design stage. The proposed framework provides a flexible tradeoff between performance and complexity. To the best of our knowledge, all existing papers considered resource allocation and beamforming design for dedicated information and energy users in a single-carrier network. In this paper, we instead build our design based on a proper nonlinear harvester model that captures the dependency of the output \gls{dc} power on both the power and shape of the input waveform, and marry the benefits of joint multi-carrier waveform and active beamforming optimization for \gls{swipt} with the passive beamforming capability of \gls{ris}, to investigate the \gls{r-e} tradeoff for one \gls{swipt} user with co-located information decoder and energy harvester. We ask ourselves the important question: \emph{How to jointly exploit the spatial domain and the frequency domain efficiently through joint waveform and beamforming design to enlarge the \gls{r-e} region of \gls{ris}-aided \gls{swipt}?} The contributions of this paper are summarized as follows.

		\emph{First,} we propose a novel \gls{ris}-aided \gls{swipt} architecture based on joint waveform, active and passive beamforming design under the diode nonlinear model \cite{Clerckx2016a}. Although this tractable harvester model accurately reveals how the input power level and waveform shape influence the output \gls{dc} power, it also introduces design challenges such as frequency coupling (i.e., components of different frequencies compensate and produce \gls{dc}), waveform coupling (i.e., different waveforms jointly contribute to \gls{dc}), and high-order objective function. To make an efficient use of the rectifier nonlinearity, we superpose a multi-carrier unmodulated power waveform (deterministic multisine) to a multi-carrier modulated information waveform and evaluate the performance under the \gls{ts} and \gls{ps} receiving modes. The proposed joint waveform, active and passive beamforming architecture exploits the rectifier nonlinearity, the channel selectivity, and a beamforming gain across frequency and spatial domains to enlarge the achievable \gls{r-e} region. This is the first paper to propose a joint waveform, active and passive beamforming architecture for \gls{ris}-aided \gls{swipt}.

		\emph{Second,} we characterize each \gls{r-e} boundary point by energy maximization under a rate constraint. The problem is solved by a \gls{bcd} algorithm based on the \gls{csit}. For active beamforming, we prove that the global optimal active information and power precoders coincide at \gls{mrt} even with rectifier nonlinearity. For passive beamforming, we propose a \gls{sca} algorithm and retrieve the \gls{ris} phase shift by eigen decomposition with optimality proof. Finally, the superposed waveform and the splitting ratio are optimized by the \gls{gp} technique. The \gls{ris} phase shift, active precoder, and waveform amplitude are updated iteratively until convergence. This is the first paper to jointly optimize waveform and active/passive beamforming in \gls{ris}-aided \gls{swipt}.

		\emph{Third,} we introduce two closed-form adaptive waveform schemes to avoid the exponential complexity of the \gls{gp} algorithm. To facilitate practical \gls{swipt} implementation, the \gls{wf} strategy for modulated waveform and the \gls{smf} strategy for multisine waveform are combined in time and power domains, respectively. The passive beamforming design is also adapted to accommodate the low-complexity waveform schemes. The proposed low-complexity \gls{bcd} algorithm achieves a good balance between performance and complexity.

		\emph{Fourth,} we provide numerical results to evaluate the proposed algorithms. It is concluded that (i) \gls{ris} enables constructive reflection and flexible subchannel design in the frequency domain that is essential for \gls{swipt} systems; (ii) \gls{ris} mainly affects the effective channel instead of the waveform design; (iii) multisine waveform is beneficial to energy transfer especially when the number of subbands is large; (iv) \gls{ts} is preferred at low \gls{snr} while \gls{ps} is preferred at high \gls{snr}; (v) there exist two optimal \gls{ris} development locations, one close to the \gls{ap} and one close to the \gls{ue}; (vi) the output \gls{snr} scales linearly with the number of transmit antennas and quadratically with the number of \gls{ris} elements; (vii) due to the rectifier nonlinearity, the output \gls{dc} scales quadratically with the number of transmit antennas and quartically with the number of \gls{ris} elements; (viii) for narrowband \gls{swipt}, the optimal active and passive beamforming for any \gls{r-e} point are also optimal for the whole \gls{r-e} region; (ix) for broadband \gls{swipt}, the optimal active and passive beamforming depend on specific \gls{r-e} point and require adaptive designs; (x) the proposed algorithms are robust to practical impairments such as inaccurate cascaded \gls{csit} and finite \gls{ris} reflection states.
	\end{subsection}
\end{section}

\begin{section}{System Model}\label{se:system_model}
	\begin{figure}[!t]
		\centering
		\def\svgwidth{0.9\columnwidth}
		\input{assets/chapter_3/system.eps_tex}
		\caption{An \gls{ris}-aided multi-carrier \gls{miso} \gls{swipt} system.}
		\label{fi:system}
	\end{figure}

	As shown in Fig.~\ref{fi:system}, we propose a \gls{ris}-aided \gls{swipt} system where an $M$-antenna \gls{ap} delivers information and power simultaneously, through an $L$-element \gls{ris}, to a single-antenna \gls{ue} over $N$ orthogonal evenly-spaced subbands. We consider a quasi-static block fading model and assume the \gls{csit} of direct and cascaded channels is known. The signals reflected by two or more times are omitted, and the noise power is assumed too small to be harvested.

	\begin{subsection}{Transmitted Signal}
		Following \cite{Clerckx2018b}, we superpose a multi-carrier modulated information-bearing waveform to a multi-carrier unmodulated power-dedicated deterministic multisine to boost the spectrum and energy efficiency. The information signal transmitted over subband $n \in \{1, \dots, N\}$ at time $t$ is
		\begin{equation}
			\boldsymbol{x}_{\mathrm{I},n}(t) = \Re\left\{\boldsymbol{w}_{\mathrm{I},n} \tilde{x}_{\mathrm{I},n}(t) e^{j2{\pi}{f_n}{t}}\right\},
		\end{equation}
		where $\boldsymbol{w}_{\mathrm{I},n} \in \mathbb{C}^{M \times 1}$ is the information precoder at subband $n$, $\tilde{x}_{\mathrm{I},n}\sim\mathcal{CN}(0,1)$ is the information symbol at subband $n$, and $f_n$ is the frequency of subband $n$. On the other hand, the power signal transmitted over subband $n$ at time $t$ is
		\begin{equation}
			\boldsymbol{x}_{\mathrm{P},n}(t) = \Re\left\{\boldsymbol{w}_{\mathrm{P},n} e^{j2{\pi}{f_n}{t}}\right\},
		\end{equation}
		where $\boldsymbol{w}_{\mathrm{P},n} \in \mathbb{C}^{M \times 1}$ is the power precoder at subband $n$. Therefore, the superposed signal transmitted over all subbands at time $t$ is
		\begin{equation}
			\boldsymbol{x}(t) = \Re{\left\{\sum_{n=1}^N(\boldsymbol{w}_{\mathrm{I},n}\tilde{x}_{\mathrm{I},n}(t)+\boldsymbol{w}_{\mathrm{P},n}){e^{j2{\pi}{f_n}{t}}}\right\}}.
		\end{equation}
		We also define $\boldsymbol{w}_{\mathrm{I/P}} \triangleq [\boldsymbol{w}_{\mathrm{I/P},1}^T,\dots,\boldsymbol{w}_{\mathrm{I/P},N}^T]^T \in \mathbb{C}^{MN \times 1}$.
	\end{subsection}

	\begin{subsection}{Reflection Pattern and Composite Channel}
		According to Green's decomposition \cite{Hansen1989}, the backscattered signal of an antenna can be decomposed into the \emph{structural mode} component and the \emph{antenna mode} component. The former is fixed and can be regarded as part of the environment multipath, while the latter is adjustable and depends on the mismatch of the antenna and load impedance. \gls{ris} element $l \in \{1, \dots, L\}$ varies its impedance $Z_l = R_l + j X_l$ to reflect the incoming signal, and the reflection coefficient is defined as
		\begin{equation}
			\phi_l = \frac{Z_l - Z_0}{Z_l + Z_0} \triangleq \eta_l e^{j\theta_l},
		\end{equation}
		where $Z_0$ is the real-valued characteristic impedance, $\eta_l \in [0, 1]$ is the reflection amplitude,\footnote{Due to the non-zero power consumption at the \gls{ris}, $R_l > 0$ in practice such that $\eta_l < 1$ and is a function of $\theta_l$. This paper sticks to the most common \gls{ris} model where the reflection amplitude equals \num{1} so as to reduce the design complexity and provide a primary benchmark for practical \gls{ris}-aided \gls{swipt}.} and $\theta_l \in [0,2\pi)$ is the phase shift. We also define $\boldsymbol{\phi} \triangleq [\phi_1, \dots, \phi_L]^H \in \mathbb{C}^{L \times 1}$ and $\boldsymbol{\Theta} \triangleq \mathrm{diag}(\phi_1, \dots, \phi_L)=\mathrm{diag}(\boldsymbol{\Phi}^*) \in \mathbb{C}^{L \times L}$ as the \gls{ris} vector and matrix, respectively.

		\begin{remark}\label{re:reflection_coefficient}
			The element impedance $Z_l$ maps to the reflection coefficient $\phi_l$ uniquely. Since the reactance $X_l$ depends on the frequency, the reflection coefficient $\phi_l$ is also a function of frequency and cannot be designed independently at different subbands. In this paper, we assume the bandwidth is small compared to the operating frequency such that the reflection coefficient of each \gls{ris} element is the same at all subbands.
		\end{remark}

		At subband $n$, we denote the \gls{ap}-\gls{ue} direct channel as $\boldsymbol{h}_{\mathrm{D},n}^H \in \mathbb{C}^{1 \times M}$, the \gls{ap}-\gls{ris} incident channel as $\boldsymbol{H}_{\mathrm{I},n} \in \mathbb{C}^{L \times M}$, and the \gls{ris}-\gls{ue} reflected channel as $\boldsymbol{h}_{\mathrm{R},n}^H \in \mathbb{C}^{1 \times L}$. The auxiliary \gls{ap}-\gls{ris}-\gls{ue} link can be modeled as a concatenation of the incident channel, the \gls{ris} reflection, and the reflected channel. Hence, the composite equivalent channel reduces to
		\begin{equation}\label{eq:h_n}
			\boldsymbol{h}_{n}^H = \boldsymbol{h}_{\mathrm{D},n}^H + \boldsymbol{h}_{\mathrm{R},n}^H \boldsymbol{\Theta} \boldsymbol{H}_{\mathrm{I},n} = \boldsymbol{h}_{\mathrm{D},n}^H + \boldsymbol{\phi}^H \boldsymbol{V}_{n},
		\end{equation}
		where we define the cascaded incident-reflected channel at subband $n$ as $\boldsymbol{V}_{n} \triangleq \mathrm{diag}(\boldsymbol{h}_{\mathrm{R},n}^H)\boldsymbol{H}_{\mathrm{I},n} \in \mathbb{C}^{L \times M}$. We also define $\boldsymbol{h} \triangleq [\boldsymbol{h}_1^T,\dots,\boldsymbol{h}_N^T]^T \in \mathbb{C}^{MN \times 1}$.

		\begin{remark}\label{re:subband_tradeoff}
			The cascaded channel varies at different subbands, but the reflection cannot be designed independently at different frequencies. Therefore, there exists a tradeoff for the passive beamforming design in the frequency domain, and the composite subchannels should be tuned adaptively to meet the specific requirement of multi-carrier \gls{swipt}. For example, one can design the reflection pattern to either enhance the strongest subband (e.g., $\max_{\boldsymbol{\phi},n} \lVert \boldsymbol{h}_n \rVert$), or improve the fairness among subbands (e.g., $\max_{\boldsymbol{\phi}} \min_n \lVert \boldsymbol{h}_n \rVert$). That is to say, \gls{ris} essentially enables a flexible subchannel design. In the \gls{miso} case, a similar effect also exists in the spatial domain. Therefore, each reflection coefficient is indeed shared by $M$ antennas over $N$ subbands.
		\end{remark}
	\end{subsection}

	\begin{subsection}{Received Signal}
		The received superposed signal at the single-antenna \gls{ue} is\footnote{We assume that the time difference of signal arrival via direct and auxiliary link is negligible compared to the symbol period.}
		\begin{equation}
			y(t) = \Re\left\{\sum_{n=1}^N{\Bigl(\boldsymbol{h}_{n}^H}{(\boldsymbol{w}_{\mathrm{I},n}\tilde{x}_{\mathrm{I},n}(t)+\boldsymbol{w}_{\mathrm{P},n})+\tilde{n}_n(t)\Bigr)}{e^{j2{\pi}{f_n}{t}}}\right\},
		\end{equation}
		where $\tilde{n}_n(t)$ is the noise at \gls{rf} band $n$. Note that the modulated component can be used for energy harvesting if necessary, but the multisine component carries no information and cannot be used for information decoding.
	\end{subsection}

	\begin{subsection}{Receiving Modes}
		\begin{figure}[!t]
			\centering
			\subfloat[\gls{ts} receiver\label{fi:ts_receiver}]{
				\resizebox{0.9\columnwidth}{!}{
					\input{assets/chapter_3/ts_receiver.tex}
				}
			}
			\\
			\subfloat[\gls{ps} receiver\label{fi:ps_receiver}]{
				\resizebox{0.9\columnwidth}{!}{
					\input{assets/chapter_3/ps_receiver.tex}
				}
			}
			\caption{Diagrams of practical co-located receivers.}
			\label{fi:receiver}
		\end{figure}

		As illustrated in Fig.~\ref{fi:receiver}, there are two practical receiving modes for the co-located information decoder and energy harvester \cite{Zhou2013}. The \gls{ts} receiver divides each transmission block into orthogonal data and energy sessions with duration $1-\eta$ and $\eta$, respectively. During each session, the transmitter optimizes the waveform for either \gls{wit} or \gls{wpt}, while the receiver activates the information decoder or the energy harvester correspondingly. The duration ratio $\eta$ controls the \gls{r-e} tradeoff and is independent from the waveform and beamforming design. On the other hand, the \gls{ps} receiver splits the incoming signal into individual data and energy streams with power ratio $1-\rho$ and $\rho$, respectively. The data stream is fed into the information decoder while the energy stream is fed into the energy harvester. During each transmission block, the superposed waveform and splitting ratio are jointly designed to achieve different \gls{r-e} tradeoffs. In the following context, we consider the optimization with the \gls{ps} receiver, as the \gls{ts} receiver can be regarded as a special case (i.e., a time sharing between $\rho=0$ and $\rho=1$).
	\end{subsection}

	\begin{subsection}{Information Decoder}
		A major benefit of the superposed waveform is that the multisine is deterministic and creates no interference to the modulated waveform \cite{Clerckx2018b}. Therefore, the achievable rate is\footnote{It requires waveform cancellation or translated demodulation \cite{Clerckx2018b}.}
		\begin{equation}\label{eq:R}
			R(\boldsymbol{\phi},\boldsymbol{w}_{\mathrm{I}},\rho) = \sum_{n=1}^N{\log_2\left(1+\frac{(1-\rho)\lvert \boldsymbol{h}_{n}^H\boldsymbol{w}_{\mathrm{I},n} \rvert^2}{\sigma_n^2}\right)},
		\end{equation}
		where $\sigma_n^2$ is the noise variance at the \gls{rf}-band and during the \gls{rf}-to-baseband conversion on tone $n$.
	\end{subsection}
